---
title: "ST443 Lab02"
date: 07 October 2020
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## lab 2.1 Install and activate packages 
```{r cars}
# install.packages("MASS")
# install.packages("ISLR")

library(MASS)
library(ISLR)
```

## lab 2.2 Simple Linear Regression

```{r}
# fix(Boston)
names(Boston)
# ?Boston
# lm.fit=lm(medv~lstat)
lm.fit = lm(medv ~ lstat, data = Boston)
attach(Boston)
lm.fit = lm(medv ~ lstat)
lm.fit
summary(lm.fit)
names(lm.fit)
lm.fit$coefficients
coef(lm.fit)
```
## lab 2.3 Inference
Confidence intervals for the coefficient estimates
```{r}
confint(lm.fit) 
```

Preditive values, confidence intervals and prediction intervals for the prediction of medv for a given value of `lstat`
$CI(\hat{y}_i)  = CI(\hat{\beta}_0 + x_i \cdot \hat{\beta}_1)$,
$PI(\hat{y}_i)  = CI(\hat{\beta}_0 + x_i \cdot \hat{\beta}_1 + e_i)$,
```{r}
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))))
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "confidence")
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "prediction")
```

## lab 2.4 Diagnostic
"NICE": normality, independent, constant variance, $E{(e)} = 0$
```{r}
plot(lstat, medv)
abline(lm.fit)
abline(lm.fit, lwd = 3)
abline(lm.fit, lwd = 3, col = "red")
plot(lstat, medv, col = "red")
plot(lstat, medv, pch = 20)
plot(lstat, medv, pch = "+")
par(mfrow = c(2, 2))
plot(lm.fit)
par(mfrow = c(1, 1))
```

plot of fitted values vs (standardized) residuals
```{r}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```

plot of leverage statistics
```{r}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

## lab 2.5 Multiple Linear Regression
```{r}
lm.fit1 = lm(medv ~ lstat + age, data = Boston)
summary(lm.fit1)
lm.fit2 = lm(medv ~ ., data = Boston)
summary(lm.fit2)
lm.fit3 = lm(medv ~ . - age, data = Boston)
summary(lm.fit3)
lm.fit4 <- lm(medv ~ lstat + age + tax + rad, data = Boston)
summary(lm.fit4)
anova(lm.fit1, lm.fit4)  ## F test, anova() function performs a hypothesis test comparing the two models
```

### Interaction term `:`.
eg.: `ptratio` and `ptratio` are not excluded from the model if using `*`. Instead, `:` should be used:
```{r}
summary(lm(medv ~ . - age + ptratio*tax, data = Boston))
summary(lm(medv ~ . - age - ptratio - tax + ptratio*tax, data = Boston)) # ptratio and tax are not excluded
summary(lm(medv ~ . - age - ptratio - tax + ptratio:tax, data = Boston)) # ptratio and tax are excluded
```

### Variables Transformation
Square transformation of `rm` is not recognized
```{r}
summary(lm(medv ~ . - age + rm ^ 2,    data = Boston))
```
You need `I()` to isolate the transformation. In other words, `I()` inhibits the interpretation of operators such as "+", "-", "*" and "^" as formula operators, so they are used as arithmetical operators:
```{r}
summary(lm(medv ~ . - age + I(rm ^ 2), data = Boston))
```

It is OK to do log transformation without `I()`:
```{r}
summary(lm(medv ~ . - age + log(rm), data = Boston)) 

```

## lab 2.6 Categorical Predictors
```{r}
# ?Carseats
names(Carseats)
```

Predict sales (child car seat sales) in 400 locations on a number of predictors
```{r}
lm.fit = lm(Sales ~ . + Income:Advertising + Price:Age, data = Carseats)
summary(lm.fit)
```

`ShelveLoc`: an indicator of the quality of the shelving location, i.e. the space within a store in which the car seat is displayed at each location
```{r}
attach(Carseats)
contrasts(ShelveLoc)
contrasts(Urban)
```

## lab 2.7 Writing Functions
```{r}
f1 = function(x){
  d = median(x) - mean(x)
  return(d)
}
f1(x = c(1,2,6))
```

```{r}
# LoadLibraries
# LoadLibraries()
LoadLibraries = function() {
  library(ISLR)
  library(MASS)
  print("The libraries have been loaded.")
}
LoadLibraries
LoadLibraries()
```