---
title: "Lab 7: Moving Beyond Linearity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ISLR)
# ?Wage
Wage = Wage
summary(Wage)
```

## Lab 6.1: Polynomial Regression and Step Functions
Fit a polynomial regression with `degree=4`. `ploy()` function generates a basis of orthogonal polynomials.
```{r}
fit <- lm(wage ~ poly(age, 4), data = Wage)
summary(fit)
```
Create a grid of values for age at which we want predictions:
```{r}
attach(Wage)
agelims <- range(age)
age.grid <- seq(from = agelims[1], to = agelims[2])
preds <-predict(fit, newdata=list(age=age.grid), se.fit = TRUE)
se.bands <-cbind(preds$fit + 2*preds$se.fit, preds$fit - 2*preds$se.fit)
```
Visualize the output:
```{r}
plot(age, wage, col="darkgrey")
title("Degree-4 Ploynomial")
lines(age.grid, preds$fit, lwd=2, col="blue")
matlines(age.grid, se.bands, lwd=1, col= "red", lty=3)
```
Some other ways of doing this in R. For example:
```{r}
fit.a <-lm(wage~ age + I(age^2)+ I(age^3)+ I(age^4), data=Wage)
summary(fit.a)
```
Two methods provide the same fitted values:
```{r}
plot(fitted(fit), fitted(fit.a))
```
Do polynomial regression with different degrees
```{r}
fit.1 <- lm(wage~ age, data=Wage)
fit.2 <- lm(wage~ poly(age, 2), data=Wage)
fit.3 <- lm(wage~ poly(age, 3), data=Wage)
fit.4 <- lm(wage~ poly(age, 4), data=Wage)
fit.5 <- lm(wage~ poly(age, 5), data=Wage)
```
Anova to compare different models
```{r}
anova(fit.1, fit.2, fit.3, fit.4, fit.5)
```
Polynomial logistic regression
```{r}
fit.logit <- glm(I(wage>250) ~ poly(age, 4), data= Wage, family="binomial")
```
(*Further details can be learnt from textbook ISLR: PAGE 292*)

Step function
```{r}
table(cut(age,4))
fit.cut <-lm(wage~ cut(age, 4), data= Wage)
summary(fit.cut)
```
***

## Lab 6.2: Splines
Fit wage to age using a regression spline
```{r}
library(splines)
fit <- lm(wage ~ bs(age, knots = c(25, 40, 60)), data = Wage)
pred <- predict(fit, newdata = list(age = age.grid), se.fit = TRUE)
plot(age, wage, col = "grey")
lines(age.grid, pred$fit, lwd = 2)
lines(age.grid, pred$fit + 2 * pred$se, lty = "dashed")
lines(age.grid, pred$fit - 2 * pred$se, lty = "dashed")
```
prespecified knots at age 25, 40 and 60 produces a spline with six basis functions (Recall that a cubic spline with three knots has *seven* degree of freedoms)
```{r}
dim(bs(age, knots=c(25, 40, 60)))
```
We could use the `df` option to produce a spline with knots at uniform quantiles of the data
```{r}
dim(bs(age, df=6))
attr(bs(age, df=6), "knots")
```
Now we fit a natural spline with four degrees of freedom
```{r}
fit2 <- lm(wage ~ ns(age, df = 4), data = Wage)
pred2 <- predict(fit2, newdata = list(age = age.grid), se.fit = TRUE)
```
To fit a smoothing spline, we use the `smooth.spline()` function
```{r}
plot(age, wage, xlim=agelims, cex=0.5, col="darkgrey")
title("Smoothing Spline")
fit<- smooth.spline(age, wage, df=16)
fit2 <- smooth.spline(age, wage, cv = TRUE)
fit2$df
lines(fit, col="red", lwd=2)
lines(fit2, col="blue", lwd=2)
legend("topright", legend=c("16 DF","6.8 DF"), col=c("red","blue"), lty=1, lwd=2, cex=0.8)
```
***

## Lab 6.3: Generalized Additive Models (GAM)

We now fit a GAM to predict wage using natural spline functions of year and age treating education as a qualitative predictor
```{r}
library(gam)
gam1 <-lm(wage ~ ns(year, 4) + ns(age, 5) + education, data=Wage)
par(mfrow=c(1,3))
plot.Gam(gam1, se=TRUE, col="red")
```

We next fit a GAM using smoothing splines rather than natural splines, here we need to use the gam library in R

`s()` function, which is part of `gam` library, is used to indicate that we would like to use a smoothing spline
```{r}
gam2 <-gam(wage~ s(year,4) + s(age,5) +education, data=Wage)
summary(gam2)
plot(gam2, se=TRUE, col="blue")
```

Since the function of year looks rather linear, we can perform a series of *ANOVA* test in order to determine which of the these models is the best

gam2: a GAM uses a spline function of year

gam3: a GAM uses a linear function of year

gam4: a GAM that excluds year
```{r}
gam3 <-gam(wage~ year + s(age,5) + education, data=Wage)
gam4 <-gam(wage~ s(age,5) + education, data=Wage)
anova(gam4, gam3, gam2, test="F")
```
we can make predictions from gam objects
```{r}
pred <-predict(gam2, newdata=Wage)
```